---
title: "Data Dictionary for ausbushfire"
author: "Heng-Hsieh Chang"
date: "2025-10-28"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Dictionary for ausbushfire}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE, message = FALSE, warning=FALSE}
library(ausbushfire)
library(dplyr)
library(knitr)
library(kableExtra)
library(tibble)
```

## Overview
The ausbushfire_data object is a primary dataset derived from the KNMI Climate Explorer and related literature, compiled to facilitate the interactive analysis of fire risk and climate drivers in Southeastern Australia (1979–2019).

## Dataset

#### Data Structure
The dataset contains annual observations for key fire weather metrics and climate teleconnections.

```{r}
# Display the structure of the dataset
dplyr::glimpse(ausbushfire_data)
```

#### Variable Definitions

This dataset uses the following variables, primarily representing annual maximum or seasonal means/totals across the study region.

```{r}
#| echo: false
table_data <- tibble::tribble(
  ~Variable, ~Units, ~Period, ~Description,
  "year", "-", "1979–2019", "Fire season year (aligned to the end of the season).",
  "fwi_sm", "-", "1979–2019", "Fire Weather Index (FWI) seasonal maximum. Primary measure of overall fire risk.",
  "msr_sm", "-", "1979–2019", "Monthly Severity Rating (MSR) seasonal maximum. **Reflects risk of prolonged, difficult-to-suppress fires.",
  "tmax", "°C", "1979–2019", "Maximum Temperature ($\\text{Tmax}$) annual mean for the fire season, a key driver of fuel dryness.",
  "precip_total", "mm", "1979–2019", "Total Precipitation during the fire season (Sep–Feb). Indicates seasonal drought conditions.",
  "iod_mean", "-", "1979–2019", "Indian Ocean Dipole (IOD) mean index (Sep–Dec). Major climate driver influencing regional drought.",
  "sam_mean", "-", "1979–2019", "Southern Annular Mode (SAM) mean index (Sep–Dec). Influences high-latitude westerly winds and rainfall patterns.",
  "area_burned", "km²", "1997–2018", "Total burned area in the study region, reflecting actual fire impact. (Data availability starts later.)"
)


kable(table_data, format = "html", escape = FALSE,
      col.names = c("Variable", "Units", "Period", "Description"),
      caption = "Field dictionary for the ausbushfire_data dataset") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed")) %>%
  column_spec(1, bold = TRUE, width = "8em") %>%
  column_spec(2, width = "5em") %>%
  column_spec(3, width = "8em") %>%
  column_spec(4, width = "40em")
```

#### Quick Peek

A look at the first few records:

```{r}
head(ausbushfire_data)
```

#### Notes:

- Data starts from 1979, but area_burned from 1997.
- Intended uses: time series trends, scatter plots for relationships, map opacity for burned area.
