---
title: "Introduction to the ausbushfire Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the ausbushfire Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ausbushfire)
```


This package provides fire weather data regarding the 2019/20 Australian bushfires from the van Oldenborgh et al. (2021) paper. It also includes an interactive Shiny app to explore this data.

## Package Data

The package contains the `ausbushfire_data` dataset. This data includes the Fire Weather Index (FWI) and Monthly Severity Rating (MSR) from 1979-2020.

You can load the package and inspect the data:

```{r setup}
library(ausbushfire)
library(dplyr)

# View the first few rows
head(ausbushfire_data)

# Get a summary of the data
summary(ausbushfire_data)
```

## Launching the Shiny App

To explore this data interactively, the package provides the `run_app()` function.

Run the following command in your R Console to launch the Shiny app:

```{r, eval=FALSE}
# Run this command to start the app
run_app()
```

The app allows you to switch between the FWI and MSR metrics and observe the time series, highlighting the extremity of the 2019/20 season.
