---
title: "index"
output: html_document
---

# ausbushfire

## Overview

The `ausbushfire` R package provides an interactive interface to explore key fire weather indices and climate drivers in Southeastern Australia from **1979 to 2019**, analyzing trends related to the Black Summer bushfires. The numbers are based on the published van Oldenborgh et al. (2021) workflow and let you:

- Browse fire weather index (FWI) and burned area on a map (with slider for year),
- Explore trends in metrics like FWI, MSR, Tmax, Precip, IOD, SAM, Area Burned (with filters),
- View relationships between metrics in scatter plots.


## Installation
You can install the development version of ausbushfire from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
# Install remotes if needed
install.packages("remotes") 

remotes::install_github("ETC5523-2025/assignment-4-packages-and-shiny-apps-Jessica19940828")

```

## Learn More

See the **Examples** → [Introduction to ausbushfire](articles/intro_to_ausbushfire.html)  
and the **Data Dictionary** → [Data Dictionary](articles/data_dictionary.html).

You can also launch the Shiny app directly with:

```{r}
ausbushfire::run_app()
```


## Sources & citation

van Oldenborgh et al. (2021) article: <https://nhess.copernicus.org/articles/21/941/2021/>

## Contributing
Issues and pull requests are welcome.
